### 2016-12-01 diagrams/context_distribution.tex
chapter 4

\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
\begin{axis}[
    ybar stacked,
	bar width=12pt,
	nodes near coords = {%
    \pgfmathprintnumberto[fixed,assume math mode=true]{\pgfplotspointmeta}{\myval}%
    \pgfmathparse{\myval<101?:\myval}\pgfmathresult%
},
    enlargelimits=0.15,
    legend style={at={(0.5,-0.28)},
      anchor=north,legend columns=-1},
    ylabel={Proportion [\%]},
    symbolic x coords={AddressBook, TimelClock, WebChess, Drupal,
    phpBB, phpMyAdmin, Anchor, Kirby, Automad, Monstra, Nibbleblog}, xtick=data,
    x tick label style={rotate=45,anchor=east},
    cycle list name=ColorListBar,
    ]

\addplot+[ybar] plot coordinates {(AddressBook, 28) 
  (TimelClock, 0) (WebChess, 86) (Drupal, 100) (phpBB, 96) (phpMyAdmin, 92)
  (Anchor, 100) (Kirby, 96) (Automad, 93) (Monstra, 83) (Nibbleblog, 91)};
\addplot+[ybar] plot coordinates {(AddressBook, 5) 
  (TimelClock, 0) (WebChess, 0) (Drupal, 0) (phpBB, 0) (phpMyAdmin, 1)
  (Anchor, 0) (Kirby, 4) (Automad, 0) (Monstra, 2) (Nibbleblog, 0)};
\addplot+[ybar] plot coordinates {(AddressBook, 67) 
  (TimelClock, 100) (WebChess, 14) (Drupal, 0) (phpBB, 4) (phpMyAdmin, 7)
  (Anchor, 0) (Kirby, 0) (Automad, 7) (Monstra, 15) (Nibbleblog, 9)};

\legend{Function Definition, HTML Template, Miscellaneous}
\end{axis}
%\caption{Relative distribution of contexts of not reached output candidates}
\end{tikzpicture}