### 2016-12-01 diagrams/coverage.tex

\pgfplotstableread[header=false, col sep=comma]{ 
	86.32, AdressBook
	100.0, SchoolMate
	92.36, TimeClock
	91.28, WebChess
	31.44, phpBB
	19.99, Drupal
	5.000, phpMyAdmin
	90.48, Anchor
	76.3,  Kirby
	46.87, Automad
	79.47, Monstra
	71.37, Nibbleblog
	}\datatablex

	\pgfplotstableread[header=false, col sep=comma]{ 
	84.44, AdressBook
	98.24, SchoolMate
	91.22, TimeClock
	88.72, WebChess
	4.76, phpBB
	18.39, Drupal
	3.79, phpMyAdmin
	89.56, Anchor
	76.30,  Kirby
	40.15, Automad
	61.38, Monstra
	59.43, Nibbleblog
	}\datatable

	\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
	  \begin{axis}[   
	  	,legend style={at={(0.25,-0.2)},
		anchor=north,legend columns=-1} 
	    ,xbar
	    ,xlabel={Coverage [\%]}
	    ,bar width=1ex, y=3ex
	    ,enlarge y limits={abs=0.75}
	    ,ytick={data} % (\empty,data,{coordinate list})
	    ,label={Reach}
	    ,yticklabels from table={\datatable}{1}
	  ]

	  \addplot[red!20!black,fill=gray!80!white] table [
	    y expr=-\coordindex, % Use negative coordinate index as y coordinate
	    x index=0 % Use first column as x coordinate
	  ] {\datatable};
	  \addplot[red!20!black,fill=RYB2!100!white] table [
	    y expr=-\coordindex, % Use negative coordinate index as y coordinate
	    x index=0 % Use first column as x coordinate
	  ] {\datatablex};

	  \legend{Output, Reached}
	  \end{axis}
	\end{tikzpicture}