### 2016-12-01 diagrams/include_coverage.tex

        \pgfplotstableread[header=false, col sep=comma]{ 
		96.77, AdressBook
		100, SchoolMate
		92.48, TimeClock
		94.64, WebChess
		65.42, phpBB
		49.03, Drupal
		39.75, phpMyAdmin
		84.38, Anchor
		39.13,  Kirby
		37.5, Automad
		37.5, Monstra
		100, Nibbleblog
		}\reachedincludes

		\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
		  \begin{axis}[   
		  	,legend style={at={(0.25,-0.2)},
			anchor=north,legend columns=-1} 
		    ,xbar
		    ,xlabel={Reached include expressions [\%]}
		    ,bar width=1.2ex, y=2ex
		    ,enlarge y limits={abs=0.75}
		    ,ytick={data} % (\empty,data,{coordinate list})
		    ,label={Reach}
		    ,yticklabels from table={\reachedincludes}{1}
		  ]

		  \addplot[red!20!black,fill=RYB10!80!white] table [
		    y expr=-\coordindex, % Use negative coordinate index as y coordinate
		    x index=0 % Use first column as x coordinate
		  ] {\reachedincludes};

		  \end{axis}
		\end{tikzpicture}