### 2016-12-01 diagrams/success_rate.tex

\pgfplotstableread[header=false, col sep=comma]{ 
		95, AdressBook
		100, SchoolMate
		100, TimeClock
		100, WebChess
		74, phpBB
		88, Drupal
		99, phpMyAdmin
		0, Anchor
		0,  Kirby
		100, Automad
		27, Monstra
		90, Nibbleblog
		}\successfulincludes

		\begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]
		  \begin{axis}[   
		  	,legend style={at={(0.25,-0.2)},
			anchor=north,legend columns=-1} 
		    ,xbar
		    ,xlabel={Successfully resolved and reached include expressions [\%]}
		    ,bar width=1.2ex, y=2ex
		    ,enlarge y limits={abs=0.75}
		    ,ytick={data} % (\empty,data,{coordinate list})
		    ,label={Reach}
		    ,yticklabels from table={\successfulincludes}{1}
		  ]

		  \addplot[red!20!black,fill=RYB11!80!white] table [
		    y expr=-\coordindex, % Use negative coordinate index as y coordinate
		    x index=0 % Use first column as x coordinate
		  ] {\successfulincludes};

		  \end{axis}
		\end{tikzpicture}