### 2017-01-07 diagrams/context_distribution.tex
Section 2 2nd iteration + back to tikz

    \pgfplotstableread{ % data 
Label     Reached/Output OnlyReached FunctionContext Bener
Adressbook 3 2 1.4 3
Schoolmate 3 2 1 2
Timeclock 3 2 1 2
Webchess 3 2 1 2
PhpBB 3 2 1 2
Drupal 3 2 1 2
phpMyAdmin 3 2 1 2
Anchor 3 2 1 2
Kirby 3 2 1 2
Automad 3 2 1 2
Monstra 3 2 1 2
Nibbleblog 3 2 1 2
    }\testdata

    \begin{tikzpicture}[thick,scale=0.8, every node/.style={transform shape}]

    \begin{axis}[
    enlargelimits=0.05,
    xbar stacked,   % Stacked horizontal bars
    xlabel={Proportion [\%]},
    xmin=0,         % Start x axis at 0
    ytick=data,     % Use as many tick labels as y coordinates
    legend style={
       at={(0.5,-0.28)}, 
       anchor=north,
       legend columns = -1,
       cells={anchor=center, fill},
       nodes={inner sep=1,below=-1.1ex}
    },
    yticklabels from table={\testdata}{Label}  % Get the labels from the Label column of the \datatable
    ,bar width=2ex, y=3ex,
    nodes near coords = {%
       \pgfmathprintnumberto[fixed,assume math mode=true]{\pgfplotspointmeta}{\myval}%
       \pgfmathparse{\myval<101?:}\pgfmathresult%
    },
    area legend
    ]
    \addplot [fill=lightgrey!80] table [x=Reached/Output, meta=Label,y expr=\coordindex] {\testdata};
    \addplot [fill=blue!60] table [x=OnlyReached, meta=Label,y expr=\coordindex] {\testdata};
    \addplot [fill=red!60] table [x=FunctionContext, meta=Label,y expr=\coordindex] {\testdata};
    \addplot [fill=red!99] table [x=Bener, meta=Label,y expr=\coordindex] {\testdata};

    \legend{Reached/Output, OnlyReached, FunctionContext, Bener}
    \end{axis}
    \end{tikzpicture}