### 2017-01-07 diagrams/coveragencontext.tex

\documentclass{article}
\usepackage{tikz,pgfplots,pgfplotstable}

  \begin{document}

    \pgfplotstableread{ % data 
Label     Adequacy Sustainability Integrity
Austria   21.87797  5.566584 24.85545
Brazil    20.93009  7.937881 24.33203
Chile     20.41148 21.126529 27.56199
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
Australia 26.83072 23.823827 28.94545
    }\testdata

    \begin{tikzpicture}

    \begin{axis}[
    xbar stacked,   % Stacked horizontal bars
    xmin=0,         % Start x axis at 0
    ytick=data,     % Use as many tick labels as y coordinates
    legend style={at={(axis cs:65,0.2)},anchor=south west},
    yticklabels from table={\testdata}{Label}  % Get the labels from the Label column of the \datatable
    ,bar width=1ex, y=3ex
    ]
    \addplot [fill=green!80] table [x=Adequacy, meta=Label,y expr=\coordindex] {\testdata};   % "First" column against the data index
    \addplot [fill=blue!60] table [x=Sustainability, meta=Label,y expr=\coordindex] {\testdata};
    \addplot [fill=red!60,
    point meta=x,
    nodes near coords,
    nodes near coords align={anchor=west},
    every node near coord/.append style={
        black,
        fill=white,
        fill opacity=0.75,
        text opacity=1,
        outer sep=\pgflinewidth % so the label fill doesn't overlap the plot
    }
    ] table [x=Integrity, meta=Label,y expr=\coordindex] {\testdata};
    \legend{Adequacy,Sustainability,Integrity}

    \end{axis}
    \end{tikzpicture}
\end{document}