### 2017-01-14 paper.tex
refactored 4 completely

\documentclass[sigconf]{acmart}

% just in case
\usepackage{etex}

% does not seem to work with sigconf template
%\usepackage{cite}

\usepackage{tikz}
\usetikzlibrary{patterns}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{listings,multicol}
\usepackage{lipsum}

\usepackage{enumerate}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{ucs}
\usepackage{framed}

\setlength\parindent{0pt}

\renewcommand{\sf}[1]{\textsf{#1}}
\renewcommand{\tt}[1]{\texttt{#1}}

\usepackage[zerostyle=d]{newtxtt}
%\renewcommand*\familydefault{\ttfamily} %% Only if the base font of the document is to be typewriter style
\usepackage[T1]{fontenc}

%\usepackage[labelfont=bf]{caption}

\definecolor{lightgrey}{HTML}{FFFFFF}

\hypersetup{
    colorlinks,
    linkcolor={red!65!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\usepackage{listings}

\lstset{
	language=PHP,
	extendedchars = \true,
	basicstyle=\ttfamily\footnotesize,
	basewidth={0.5em,0.4em}, % for other columns settings
	commentstyle = \color{gray},
	columns=fullflexible,
	numbers=left,
	numberstyle=\tiny,
	frame=tb,
	showstringspaces=false,
	xleftmargin=2em,
	framexleftmargin=1.5em,
	backgroundcolor=\color{lightgrey},
	inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries
}

%\newcommand{\dollar}{\mbox{\textdollar}}

\newcommand*{\origrightarrow}{}
\let\oldarrow\textrightarrow
\renewcommand*{\textrightarrow}{\fontfamily{cmr}\selectfont\origrightarrow}

% ARIAL
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}

% TIMES
%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx

\usepackage{pgfplots, pgfplotstable}

\usepackage{xcolor}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[ISSTA'17]{ACM Woodstock conference}{July 2017}{Santa Barbara,
California, USA}
\acmYear{1997}
\copyrightyear{2016}

\acmPrice{15.00}

\begin{document}
\title{Does It Scale? Static Output Approximation of PHP Web Applications Using
Symbolic Execution}
\titlenote{Produces the permission block, and copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

\author{Stefan Mühlbauer}
\affiliation{%
  \institution{Technische Universität Braunschweig}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

\author{Christian Kästner}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{Carnegie Mellon University}
  %\streetaddress{5000 Forbes Avenue}
  %\city{Pittsburgh} 
  %\state{PA} 
  %\postcode{15217}
}
%\email{s.muehlbauer@tu-bs.de}

\author{Tien N. Nguyen}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{University of Texas at Dallas}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{\textcolor{blue}{put shortauthors here}}

\begin{abstract}
Dynamic web applications have become widely popular and are to
a large proportion based on the scripting language PHP. Output approximation of web
applications enables a range of additional tool support as well as possibilities
for vulnerability detection. Unfortunately, recent approximation approaches have
only been evaluated for smaller systems.

This paper presents an experience report about the scalability of output
approximation using symbolic execution of state-of-the-art PHP web
applications. For a symbolic execution engine extended with support for
object-oriented programming and arrays, we identified language features and
corresponding programming patterns that impede symbolic execution and limit the
scalability of this approach. Our findings include: (1) Dynamic features such as
functions and includes are prone to fail for certain programming patterns. (2)
Expressions containing elements from I/O, databases or files can heavily impede
symbolic execution. Our findings provide useful guidelines to design new
tools and also to improve the development process of statically analyzable web
applications.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

% We no longer use \terms command
%\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\section{Introduction}
With the emerging world wide web, dynamic web applications have become
widely popular. Various different implementation techniques have emerged,
ranging from technologies for programming languages (JSP, ASP .NET) and web
application frameworks for  script languages (Ruby On Rails, Django) to
languages tailored specifically to the domain of web applications. PHP
cite{phpNET} is a programming language focused on server-side application
development. As of 2012, it was used by 78.8 percent of the ten million most popular websites (according
to Alexa popularity ranking) \cite{alexaPHP}, ranking 7th on the TIOBE
programming community index \cite{tiobePHP}, and was the ranked as the 6th
most popular language on GitHub \cite{githubPHP}.

One common property of all technologies for dynamic web applications is
\emph{staged computation}: A dynamic web application as a whole
consists of both static code, such as scripts, and dynamically generated code,
such as client page output. The latter code, although assembled at
runtime, may contain client-side parts of the web application, such as
JavaScript. So, to study dynamic web applications in its entirety, we need to
consider both static as well as dynamic aspects of systems.

\emph{Symbolic execution} is a static analysis technique between program testing
and program proving. For a program, symbolic values are
used instead of concrete inputs \cite{Darringer1978,King1976}. The underlying
concept is to map program input to program output: Symbolic values are used and
propagated throughout the symbolic execution and keep dependencies for program
output traceable. This analysis technique helps unfold the \emph{staged} nature
of dynamic web applications. Since every feasible path can be executed the
symbolic output contains both invariant output as well as output variants that
depend on program input.

Knowledge about different output output variants is leveraged by a number of
analyses for the domain of web applications. Previous work presented tools
for detecting and locating HTML validation errors
\cite{Nguyen:2011:AFH:2190078.2190142}, computing program slices across
server-side and client-page code \cite{Nguyen:2015:CPS:2786805.2786872}, or
easing development and maintenance by extending IDE support for web applications
with code navigation
\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.

Despite the various use cases for analyses based on an approximated output
model, static output approximation for PHP web applications using symbolic
execution has so far only been evaluated for small systems that are not
maintained anymore. The tools presented by previous work
\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2011:AFH:2190078.2190142}
though are only practical, if for a given system the symbolic execution engine
is scalable, i.e., will also approximate output accurately for larger and more
recent systems with acceptable time and space consumption.

To investigate the question, whether we can have a practical and scalable
symbolic executoin engine, we re-implemented the engine with the specifications
of the previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
as object-oriented programming. We evaluate our
symbolic execution engine for large-scale and modern PHP systems.

During the introduction of new semantics, we addressed two  trade-offs between
accuracy of our e xecution and performance: For method calls with ambiguous
targets as well as concrete execution of loops the number of program states can
become infeasible. This may require additional engineering adjustments to tame
the state space explosion at cost of accuracy.

Based on out observations we identify conceptual limitations of symbolic
execution for PHP.  Dynamic features such as indirect function calls and
include expression require concrete information to be evaluated properly. Since
these expressions may be assembled at runtime and can contain symbolic
information, for many cases this restricts our engine from further execution.
Although some effort to statically approximate include expression can be spent,
it is often non-trivial since expressions can contain information coming from
various inputs including databases, user inputs or configuration files. We
empirically evaluate our observations for a corpus including large PHP systems that are still maintained.

Our key contribution in this paper include (1)  a new tool infrastructure to
statically approximate the client-page output for PHP web applications using
symbolic execution and (2) a report of out observations for state-of-the-art
web applications as well as an empirical evaluation of conceptual limitations
of symbolic executions for PHP web applications.

\section{State Of the Art}\label{sec:stateoftheart}
Symbolic execution is an analysis techniques used by the tools
whose practicality we aim to evaluate. This section recaps the idea of symbolic
execution, the tools which make use of it, and both conceptual and
language-related limitations.

\subsection{Symbolic Execution}
Symbolic execution is a static analysis technique that was proposed by James C.
King in 1976 \cite{King1976}. Symbolic execution allows to explore feasible
paths in a program. Based on normal program execution, the execution semantics
is extended to handle symbolic values. An execution becomes symbolic by
assuming symbolic values as program inputs rather than obtaining concrete
values \cite{King1976,Darringer1978}. These symbolic values are then propagated
and used throughout the execution.

Execution starts with a tautology path
condition. The control-flow can be split whenever the decision between branches
is ambiguous. This is the case, when conditional expressions, for example for
if-statements, evaluate to a symbolic value rather than an actual result.
The corresponding path conditions (guards for instance) are conjoined with the
previous path condition \cite{King1976}. Additionally, the conjoined path
condition can be checked for satisfiability in order to exclude infeasible
paths.

\subsection{Static Output Approximation for PHP} \label{sec:workflow}
Symbolic execution is not dedicated to a particular
programming language. For dynamic web applications, in our case for web
applications written in PHP, it can be used to approximate all possible client
page output variants. Before we present in detail existing tools that aim to
ease development and maintenance of PHP web applications, we illustrate how output
approximation using symbolic execution is pursued by starting with a small
example.

\begin{figure}[t]
	\centering
	\begin{subfigure}[center]{0.48\textwidth}%%[multicols=1]
\begin{lstlisting}
<form>
  <input type="text" name="name">
  <input type="submit" value="Submit">
</form>

<?php
function make_titles($arg, $n) {
  $is = shuffle(array(1,2,3));
  while ($n < $is[0]) {
    echo "<h1>" . $arg . "</h1><br />";
    $n = $n + 1;
  }
}

call_user_func("make_titles", "Headline", 4);

$greeting = "Hello " . $_POST['name'] .  "!";
if (!isset($_POST['name'])) {
  echo "No name entered!";
} else {
  echo $greeting;
}
?>
\end{lstlisting}
		\caption{Snippet of HTML and PHP server-side web application code.
		\label{fig:web_application}}
    \end{subfigure}

	\vspace{2mm}

    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[center]{0.48\textwidth}
\begin{lstlisting}[mathescape]
<form>
  <input type="text" name="name">
  <input type="submit" value="Submit">
</form>

// #repeat n < shuffle(is)
<h1>Headline</h1><br />
// #endrepeat

// #if !isset(name)
No name entered!
// #else
Entered $\alpha$!
// #endif

\end{lstlisting}
		\caption{
			Approximation client paege output of the server-side web application code in
			Figure \ref{fig:web_application}. Variability as well as repetition annotations
			are represented by preprocessor directives.
			\label{fig:output_example}
		}
    \end{subfigure}
    \vspace{2mm}
    \caption{
    	Illustration of static ouptut approximation using symbolic
    	execution
    	\label{fig:code_example}
    }
\end{figure}

Figure \ref{fig:code_example} shows two code listings. The snippet of server-side web
application code in Figure \ref{fig:web_application} contains both HTML code
(input form in lines 1 -- 4) as well as PHP code: First, the function
\texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} passed as an argument between zero and an arbitrary number of
times. Second, the function is indirectly called using the built-in mechnism
\texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
\texttt{4} as arguments.  Finally, the input from the form in lines 1 – 4 is
used: The variable \texttt{\$greeting} consists of three concatenated strings,
\texttt{Hello}, the value of input field \texttt{name}, and an exclamation mark.
For the greeting there are two possible client page output
variants: If the input field \texttt{name} is empty at runtime, line 19 is
executed and \texttt{No name entered!} is printed, or, if a name was entered,
line 21 is executed and the greeting constructed in line 17 is printed.

The following symbolic execution semantics for PHP features based on
previous work \cite{Nguyen:2014:BCG:2635868.2635928} illustrate the necessary
modifications and extensions made to handle the level of ambiguity that comes with handling symbolic values. In addition to normal output, the
output approximation contains preprocessor directives to express
repeatable parts as well as output variants. All approximated client page
output corresponding to Figure \ref{fig:web_application} consisting of two different variants
is illustrated in Figure \ref{fig:output_example}.
First, for loops in the previous execution semantics
\cite{Nguyen:2014:BCG:2635868.2635928} only one iteration is executed instead of
an arbitrary number of iterations. Since a loop condition can contain symbolic
values, it might not be feasible to determine whether, and if so, when a loop
execution terminates. Hence, any output constructed during the single loop
iteration is highlighted (\texttt{repeat \ldots endrepeat}) to possibly be repeated an arbitrary number of times. The output in line 7 is marked to be repeated an arbitrary number of times, as the number of loop iterations cannot be determined statically. Second, a similar procedure
applies to recursive functions. Once called, a recursive function is  also only
executed to recursion depth one per call, and any subsequent invocation of that
function returns a symbolic value. This, again, is due to the difficulty to
determine the exact recursion depth similar to the number of loop iterations.
Third, any form of user interaction providing input data as well as any form of input referred
to from the deployment context of the web application, i.e. configuration files
or databases, is assumed to be symbolic. The context variable
\texttt{\$\_POST} represents input submitted to the server and, hence, is
symbolic as every program input is per definition substituted by symbolic values. Finally, the
output in lines 11 and 13 respectively are marked as alternative variants
(\texttt{if \ldots else \ldots endif}) as symbolic execution of lines 18 – 22
in Figure \ref{fig:web_application} will explore both different branches.

\subsection{Existing Tool Support for PHP}\label{sec:existing_tools}
All tools proposed so far leverage a representation of all possible HTML client
page output. Since any single client page output variant can be analyzed this
enables tool support addressing the web application as a whole in spite of its
staged nature, where parts of the web application are
generated dynamically.
Based on the approximated output representation, subsequent analyses can be
conducted:
\begin{itemize}
	\item \emph{PhpSync}: Using all variants of  HTML client page output, every
	single one can be statically checked for markup validity, i.e., if the web page
	conforms to syntactical specifications for HTML and other client-side
	languages. After the tool detects validation errors, auto-fixes
	can be provided. Otherwise, it traces validation errors back to
	source code responsible for the defect, which then can be refactored manually
	\cite{Nguyen:2011:AFH:2190078.2190142}.

	\item \emph{WebSlice}: Program slices enable to extract and understand the
	impact of changes in an application. To consider all of a dynamic web
	application for program slices, client page output needs to be taken into
	account. WebSlice combines PHP data-flow information with an output
	approximation to enable program slices across different languages
	\cite{Nguyen:2015:CPS:2786805.2786872}.

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	ouput approximation a callgraph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.
\end{itemize}

\subsection{Assumptions And Limitations}\label{sec:limitations}
In addition to the assumptions introduced by symbolic execution, such as
symbolic inputs and one-time loop execution, the symbolic execution
engine described in previous work \cite{Nguyen:2014:BCG:2635868.2635928} does
not support advanced language features, including object-oriented programming. Moreover, the range of
standard library functions supported are mostly string operations, yet PHP
provides a large collection of functions for array operations.
All three analyses described in Section \ref{sec:existing_tools} though have
only been applied to systems that are small and not maintained any more. Despite the
functional benefits these tools provide to developers, their practicality for
large and modern systems with respect to advanced language support has not been
evaluated.

\section{Scalable Output Approximation}%
The main objective of this experience report is to investigate whether we can
have a scalable symbolic execution engine as it is the underlying technique for
static output approximation in previous work
\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}.
Our vision is a practical symbolic execution, i.e., it computes an accurate approximation for large and modern PHP systems with reasonable time and space consumption.

Also, we are interested in the limitations of scalable symbolic execution for
PHP web applications. Limitations may either be conceptional, where symbolic
execution of PHP code, in contrast to concrete execution, is inaccurate: For
example an include expression evaluated to a symbolic value simply provides to
little information to include a script. Or, in spite of infeasible effort
required, limitations could allow an accurate symbolic execution. For example,
exhaustively exploring all feasible paths in an application may result in an
infeasible number of program states to store; nevertheless, this so called
state explosion problem may be addressed with additional engineering.

\subsection{Towards Scalable Output Approximation?}
As our goal is to have a scalable symbolic execution engine, we approach this
question by re-implementing the symbolic execution semantics described in
previous work \cite{Nguyen:2014:BCG:2635868.2635928} and extending it with
support for additional language features. We aim to support symbolic execution
for a real-world example system of reasonable size that incorporates the
missing language features described in Section \ref{sec:limitations} and is
still developed and maintained. \textsf{WordPress} is a popular open-source
Content Management System (CMS) with around \textcolor{blue}{300.000} lines of
code that provides a vivid plug-in environment and is maintained by a large
community.
For our implementation of a symbolic execution engine for PHP,
\textsf{Oak}\footnote{The implementation can be found
at \url{github.com/smba/oak}.}, we have chosen a test-driven approach with
continuous integration using regression tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section \ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for
object-oriented orientation (class definition, class instantiation and method
invocation) as well as most PHP array operations. A list of the functions
provided by the PHP standard library can be found at \cite{PHP}.
Throughout the implementation we did not achieve high code coverage for
\sf{WordPress}. In the following we present our observations we made thorughout
the implementation regarding issues impeding symbolic execution.

\subsection{Experience Report}\label{sec:experience_report}
We identified trade-offs between accuracy and performance that may
cause the approximation to be impractical, yet possible. Additionally, we
detected limitations impeding symbolic execution conceptually: For certain
language features, concrete information rather than symbolic values may be required to
actually be executed properly as this defective execution can impact subsequent
parts of the program which may become inaccessible and may be missed.

\subsubsection{Symbolic vs Concrete Execution}
Throughout the implementation, we basically followed the symbolic execution
semantics described in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.
Nevertheless, for additional language features we experienced it to be beneficial to the
approximation to execute them as precisely as possible.
For instance, \sf{WordPress} uses a text filter mechanism, where different
filter functions are applied sequentially to a string literal. Filters can
be linked to a hook using  associative arrays so that when the hook is
triggered, all filters that are linked to it are applied. As described in
Section \ref{} we execute loops only once due to an possibly
indeterminable number of iterations. PHP yet offers more loop constructs: the
do-while loop and the foreach loop. While for the do-while loop as well as for
the while loop the number of iterations is possibly indeterminable, the foreach
loop iterates over all items of a given array whose number of items is known at
runtime. For \sf{WordPress}, we are able to achieve a more accurate
approximation and higher code coverage results by concretely executing foreach
loops as more server-side code (filter functions in our case) are accessed and
executed. In turn, exhaustively executing foreach loops and filter functions
concretely is a trade-off between accuracy and performance, as a single hook
call site can be called from various contexts since we are exploring multiple
execution paths.

An additional aspect on the trade-off between concrete and symbolic execution,
or effort and accuracy respectively, is the extent to which program inputs are
assumed to be symbolic. As shown in Section \ref{} introducing symbolic values
can be problematic if they are propagated to language features requiring
concrete information, such as include expressions or indirect function calls.
We manually scaled the number of sources of symbolic inputs by providing static
dummy inputs such as a concrete form input, file name or configuration entry.
For specific cases, where concrete information is inevitably required, we can
achieve higher code coverage and a more accurate approximation, as more code
becomes accessible. Nevertheless, while detecting dynamic features can be
automated, it requires manual inspection and understanding of the web
application to provide appropriate dummy input data. Again, this is a trade-off
between the effort to provide input data and accuracy of the approximation and
code coverage respectively.

\subsubsection{Multi-Target Method Calls}
A non-conceptual problem we encountere is related to object-oriented programming. While exploring
different execution paths when the control flow is split, a variable can have
different values depending on the path condition under which a value has been
assigned to it (see the code example in Figure \ref{fig:output_example}).
Similarly, a variable can point to different objects depending on the path
condition as the snippet in Figure \ref{fig:multi_target} illustrates. The variable \tt{\$person} can either
point to object \tt{Alice} or \tt{Bob}. In addition, for any method invocated on
that variable the actual target is ambiguous as there are two, depending on the
path condition. In order to have a sound symbolic execution, both method calls
need to be executed. Nevertheless, the number of contexts for which a method can
be called can grow rapidly, and easily become infeasible to execute
exhaustively.

We implemented several heuristic approaches to tame this state space explosion
for multi-target method calls, including executing a method only for a subset of
target objects, or just a single object. This is a trade-off between accuracy
and soundness of the symbolic execution and performance time required.

\subsubsection{Dynamic Language Features}
\label{sec:experiment_dynamicfeaturess} 
Conceptually, we identified two language features for which symbolic execution
failed frequently. As stated in Section \ref{sec:workflow}, dynamic web
applications are staged and client page output is computed at runtime. Symbolic
execution has shown to be a useful technique to approximate dynamically
generated output in theory, though it is challenged by language features that
require concrete information rather than symbolic information at runtime. In particular, this
applies to language features that provide access to reusable code of the web
application: include expressions and function calls.

First, include expressions in PHP usually require a string value representing
the path to the script file or template to include. If this string value is
statically provided by a string literal or can be constructed unambiguously, the
symbolic execution engine was able to resolve the include expression and include
the desired file. However, it is also common to assemble include expressions not
only from static string literals, but from several sources including user input
information, database query results or configuration files. Figure
\ref{fig:dynamic_includes} shows a small example of a dynamic include containing
information retrieved from a database. Since any input for a symbolically
executed application is symbolic this information can be propagated to include
expressions. Also, configuration files represent only one particular
configuration or even just default values or placeholders. Hence, symbolic
execuion is prone to fail for dynamic includes if symbolic information is contained.

Second, function can either be called directly by statically providing the name
of the function (or method), or indirectly. Indirect function calls in PHP are
enabled by using built-in functions like \tt{call\_user\_func},
\tt{call\_user\_func\_array} or simply passing a string value with a direct
function call to the \tt{eval} function. The first two functions take as
arguments the name of the function two call, and all arguments to that
functions either as additional arguments or a single associative array
respectively.
The \tt{eval} function evaluates a given PHP expression passed as a string
literal.
Again, we encountered issues similar to dynamic include expressions with
indirect function call mechanisms: The example in Figure
\ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
\tt{header\_sserif} which both print a HTML headline. Depending on whether the
value of the variable \tt{\$style} retrieved from a database is \tt{serif} or
\tt{sserif}, a different function is called. Given that the function name is 
concrete at runtime, the desired function can be called. Otherwise, if the
function name is assembled at runtime and contains symbolic values, no function
can be called as the name is ambiguous or unknown. For this example, symbolic
execution is not able to determine which function to call indirectly.

% code examples for 3.3 experience report
\begin{figure}[t]
	\begin{subfigure}[b]{0.48\textwidth}

\begin{lstlisting}
<?php
class Person {
   function __construct__($name) {
      $this->name = $name;
   }
   function greet($greeting) {
      echo $greeting . $this->name;
   }
}

if (...) {
   $person = new Person("Alice");
} else {
   $person = new Person("Bob");
}
$person->greet("Good Morning, ");
\end{lstlisting}
	\caption{
		Dynamic dispatch: Multiple targets for method calls
		\label{fig:multi_target}
	}
	\end{subfigure}
	\vspace{2mm}
	\hfill
	\begin{subfigure}[b]{0.48\textwidth}

\begin{lstlisting}
<?php
define("ROOT", getcwd());
define("TEMPLATES", ROOT . "templates/")

$template = mysql_result(...);

require_once TEMPLATE . $template;
\end{lstlisting}
	\caption{
		Dynamic Include Resolution
		\label{fig:dynamic_includes}
	}
	\end{subfigure}
	\vspace{2mm}
	\hfill
	\begin{subfigure}[b]{0.48\textwidth}

\begin{lstlisting}
<?php
function header_serif($title) {
   $style = "font-family: serif;";
   echo "<h1 style = '$style'>$title</h1>";
}
function header_sserif($title) {
   $style = "font-family: sans-serif;";
   echo "<h1 style = '$style'>$title</h1>";
}

$font_style = mysql_result(...);
call_user_func('header_' . $font_style, 'Title');
\end{lstlisting}
	\caption{
		Function calls by indirect invocation mechanisms
		\label{fig:indirect_calls}
	}

	\end{subfigure}
	\caption{
		Code Examples of defective code features
		\label{fig:defective_code_features}
	}
\end{figure}

\subsubsection{Interpreter Customization}
We have approached issues we encountered, including
multiple-target method calls and foreach loops, in different ways as these
represent a trade-off between accuracy of execution and performance.
For method calls having ambiguous targets, we implemented different modes of
execution, ranging from a complete mode where a method is executed for all
possible targets to abstract modes with either a subset of targets or only a
single target. In the latter cases where targets are discarded, we assumed
symbolic values as method return values. Our symbolic interpreter can be
customized to work in the desired mode to scale accuracy or performance.
Nevertheless, for \sf{WordPress} these variations did not have significant
impact on the output approximation as symbolic execution was rather impeded by
dynamic features.
Also, for foreach loops we are able to toggle between a single iteration mode
and a complete iteration mode. Although we were able to increase code coverage
by accessing more code, it did not significantly increase our output
approximation.
Note that these variants have only been tested for \sf{WordPress} to better
understand possible trade-offs or simply achieve any output approximation in a
reasonable period of time. We believe based on our experience dynamic features
to be the main conceptual limitation impeding code coverage as well as output
approximation quality. In the next Section we analyze in detail the impact of
dynamic features with concrete foreach loop execution and complete
multiple-target execution for method calls.

\section{Evaluation}
As already stated in the previous Section \ref{sec:experience_report}, we
encountered several problems during the implementation of our symbolic
execution engine for \sf{WordPress}. The main items of critique of the symbolic
execution engine and tools described in previous work
\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
are the missing support for widely-used language features, and the evaluation
only using small-scale systems. Although evaluating a tool for a small systems
can be a valuable proof-of-concept, it does not necessarily cover the question
of whether a tool is practical on a larger-scale. Therefore in the following we
present our methods to evaluate practicality of our symbolic output
approximation. Moreover, we investigate to which extent the conceptual limitations encountered
with \sf{WordPress} apply for a wider choice of PHP systems.

\subsection{Experiment Setup} \label{sec:experiment_setup}
We have encountered several issues for \sf{WordPress} that conceptually limit
the scalability of symbolic execution. Since our observations so far only considers
one example system an, we evaluate our symbolic execution engine for a wider
range of PHP systems with regard to the following questions. The metrics to
approach those questions are explained in Sections
\ref{HowAccurateIsOurApproximation} and \ref{sec:understanding_limitations}.

First, the evaluation of the output approximation using the symbolic execution
engine, \sf{Symex}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928} has
shown high code coverage for small systems. As our symbolic execution
re-implements \sf{Symex} and extends with support for additional language
features, we ask whether we can replicate high coverage results at least on a small-scale. Therefore, our PHP
corpus contains four small-scale systems used in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.

Second, we ask whether system size and usage of modern language features are
factors affecting the practicality of our symbolic execution engine. We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
footnote}, did not support a number of language features. The full list of PHP
systems for that we evaluate our symbolic execution engine along with
descriptive code metrics can be found in Table \ref{corpus}.

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{4cm}rrrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}\\
	\midrule
	AddressBook & 8.2.5.2 &  & 51,907 & 239  & 1009 & 186\\
	SchoolMate & 1.5.4 &  & 8,118 & 65  & 853 & 88\\
	TimeClock & 1.04 &  & 20,800 & 63  & 7920 & 306\\
	WebChess & 1.0.0 &  & 5,219 & 28  & 470 & 56\\
	\midrule
	Drupal & 7.5.0 & CMS & 52,464 & 125  & 3569 & 749\\
	phpBB & 3.1.9 &  & 327,371 & 1,398  & 3606 & 206\\
	phpMyAdmin & 4.6.3 &  & 303,582 & 871  & 7103 & 571\\
	\midrule
	Anchor & 0.12.1 &  & 15,054 & 201 & 987 & 32\\
	Kirby & XXX & CMS & XXX & XXX  & 654 & 23\\
	Automad & XXX & CMS & XXX & XXX  & 655 & 8\\
	Monstra & XXX & CMS & XXX & XXX  & 1934 & 48\\
	Nibbleblog & XXX & CMS & XXX & XXX  & 1013 & 28\\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

\subsection{Measuring Approximation Success} \label{heuristic}
In order to accurately answer the question how good our symbolic execution
engine approximates client page output, we would require knowledge of all
possible variants to compare our approximation against. Since we do not have
any ground truth information alike, we approach this demand by using a
heuristic. Rather than having knowledge about all client page output variants
we say our approximation is accurate if all string literals (which are embedded
in and  scattered across server-side code) that may eventually become part of any
client page output variant (in the following referred to as \emph{output
candidate}) is contained in our symbolic output approximation.
It is non-trivial to determine whether a string literal is an output candidate,
so we heuristically classify string literals as output candidates if they
contain the characters \tt{<}, \tt{>}, or both since we expect output to contain
HTML tags.

We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from our corpus of PHP systems (see Table
\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 96 percent of the string literals are classified correctly. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output candidates, actually were responsive to the classifier. We
attempted to increase recall by looking for further distinctive properties to build a
classifier from, but could not do so without decreasing precision. 
\textcolor{blue}{In spite of missing half of the output candidates, we decided to use a simple, yet
precise and distinctive classifier.}

\subsection{Measuring Approximation Accuracy}
\label{HowAccurateIsOurApproximation} 
In Section \ref{heuristic} we introduced the definition of output candidates as
expected output. For an approximation to be most accurate, it needs to contain
all output candidates of a system analyzed. Based on whether output
candidates are reached by an execution path and part of the output
approximatiopn define two metrics to approach measuring accuracy.

We measure how much of the expected output was actually processed by the
symbolic interpreter.  Once a line of code, a statement or expression
containing an output candidate is actually an element in an execution path, we
define this output candidate as \emph{reached}. We define the metric \emph{reach
coverage} as the ratio of output candidates that are reached and the total
number of output candidates in the analyzed system. Although a reached output
candidate is processed by the symbolic interpreter, it does not guarantee we
will see that output candidate in the symbolic output. Therefore, we define the
metric \emph{output coverage} as the ratio of output candidates that are
contained in the output of the symbolic interpreter and the total number of
output candidates in the analyzed system.  
%For an imprecise approximation, we may see loss of output candidate information
% resulting in a output coverage lower than the reach coverage.

The coverage results for our case study are illustrated in Figure
\ref{coverage}. We could replicate high coverage results for the four small-scale systems with a reach
coverage and output coverage over 80 percent respectively. For the three
modern/large-scale systems we measured poor coverage with reach coverage
ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
We measured medium to high coverage for the more recent small-scale systems:
reach coverage ranging from 43 to 85 percent, output coverage ranging from 40
to 85 percent.

\subsection{Understanding Limitations} \label{sec:understanding_limitations}
As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
coverage metrics were poor for large-scale and small-scale/recent systems. To
understand what output candidates we missed and why, we further investigated
our approximation results and conducted three more measurements.

% RESULT DIAGRAMS
\begin{figure}[h!]
	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/not_reached}
		\caption{\label{coverage}}
	\end{subfigure}

	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/includes}
		\caption{
			Include coverage and accuracy.
			\label{fig:include_coverage_results}
		}
	\end{subfigure}

	% Callback candidate explanation
	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/callbacks}
		\caption{
			Distribution dead function output candidates that are par-
			tially or completely explained due to callback candidates.
			\label{fig:output_candidate_explanation}
		}

	\end{subfigure}
	\caption{Results}
\end{figure}

\subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out
whether their surrounding program code was not accessed/accessible, and if so,
why. We are interested in those output candidates that are located in HTML
files (possibly with nested PHP), or located in a function definition. Plain
HTML files do not require any further execution and represent output candidates
that just need to be included properly to be part of the output. Aside, output
candidates that are part of a function definition are only missed if the
corresponding function is never called at runtime.

We started classifying missed output candidates by their string literal
context. As the classification statistics in Figure \ref{coverage} illustrate,
for almost all systems tested (except for \textsf{AddressBook} and
\textsf{TimeClock}) not-reached output candidates had a function context.

%Note that \textsf{SchoolMate} was excluded from the diagram since our analysis
% reached all output candidates for that system.

%\subsubsection{Inaccessible Dynamic Features} \label{sec:inaccessible}
For both cases of context, either inclusion of a script file failed, or a
function call failed. If an HTML file is not part of the output, it also is
never reached, i.e., included. An include can fail due to various reasons, such
as an imprecise evaluation of the include expression returning a symbolic
value, or simply a missing file.

For a function to be never reached there are several scenarios: A function is
undefined at runtime if the corresponding script file is never included. In
turn, if the function is defined at runtime, the function is dead
code if there is no call site for this function, or the function is only called
indirectly and the function name could not be resolved accurately. In PHP there
are several ways to call a function beside direct call sites. 
%The language offers indirect call mechanisms like callback-commands or an
% evaluation function that parses and evaluates PHP source code as strings.

Given a symbolic value, indirect call mechanisms like callback-commands are
likely to fail since target function name and the symbolic value do not match.
This also applies for the evaluation of include expressions as concrete values
representing include targets can be included, but any include target containing
symbolic values is ambiguous. Since it is unlikely that all functions containing
missed output candidates are dead code having no call site (direct or
indirect) at all, in Section \ref{WhyDidFunctionCallsFail} we further
investigate how many functions fail due to inaccurately resolved indirect
function calls. Moreover, we evaluate the accuracy of evaluation of include
expressions in the next section.

\subsubsection{Why did includes fail?}
\label{WhyDidIncludesFail}
From Section \ref{sec:experiment_dynamicfeaturess} we learn that include
expressions can not be evaluated accurately for symbolic values, and from the
previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed output
candidates can be explained by simply missing to include a file. To better
understand the impact of include expression resolution throughout analyzing
systems, we construct two additional metrics. First, we are interested in how
many include expressions throughout the execution we actually reach in a system.
Therefore we introduce the metric \emph{reach coverage for include expressions}
as the ratio of reached include expressions and the total number of reach
expressions in a system to check whether failed includes may have caused
subsequent include expressions and so on. Second, we measure how successful
include expressions are resolved.

Figure \ref{fig:include_coverage_results} illustrates the classification of
include expressions for each system: We can either reach an include expression
and successfully resolve it, or not. Or, we did not reach an include expression at all. 
Figure \ref{fig:include_coverage_results} indicates that except for the
small-scale systems as well as \sf{Anchor} and \sf{Nibbleblog} we did not really
reach a great portion of include expressions, although for most systems the
resolution success rate was over 80 percent. Note that one failed attempt
to include a script file can result in a cascade of missing more include
expressions as the include target script file is not included and executed.

\subsubsection{Why did function calls fail?}
\label{WhyDidFunctionCallsFail}
Aside from failed or successful includes, this metric describes one scenario of
why functions containing not-reached output candidates are never called: The
trivial explanation would be that there are no direct calls of those functions,
yet PHP offers a number of ways to call a function indirectly, for instance
using built-in functions, or simply evaluating strings as PHP code using the
\texttt{eval} function.

Moreover we need to take into account that one missed function execution can
result in missing even more function calls (direct or indirect) and executions.
Thus, for functions containing not-reached output candidates we measure whether
they can be (1) only, (2) partially or (3) never accessed transitively from
functions that have no direct call sites. We refer to those functions (or
access points) having no direct call site as callback candidates since the only
way to reach them is through indirect mechanisms. Measuring how many
not-reached output candidates can be explained by functions that are callback
candidates helps to understand whether, and if so, indirect call mechanisms and
their usage may impact our analysis’ code coverage.

The callgraph snippets in Figure \ref{fig:callgraph} illustrate the idea of
detecting the functions’ access points: The nodes at the top depict callback
candidates, the nodes at the bottom with labels \tt{X}, \tt{Y}, and \tt{Z}
depict function definitions containing not-reached output candidates. In Figure
\ref{fig:callgraph} function \tt{X} can only be accessed through callback
candidates, as for Figure \ref{fig:callgraph} function \tt{Y} can be partially
accessed by callback candidates, whereas function \tt{Z} is never accessible
through callback candidates.

% Callback candidate description
\begin{figure}[t!]
	%\centering
	\includegraphics[scale=0.55]{images-paper/callgraph_cases}
    \caption{Callgraph analysis: The bottom nodes represent functions
    containing missed output candidates, the top nodes depict entry points to
    them.}
    \label{fig:callgraph} 
\end{figure}

%Like for include expressions, we again sampled dead function cases and
% identified the lack of direct call sites to those functions. As described in
% Section \ref{sec:inaccessible}, functions despite having no direct call sites
% were called through indirect call features. We identified dynamically
% assembled expressions (function names) to be responsible for failed attempts
% of calling a function indirectly as the resolved function name eventually
% contained symbolic values. These expressions, as well as include expressions,
% contained information dependent on the system environment.

For failed function calls to functions containing not-reached output
candidates we measured how many of those functions were only, or
partially accessibly through indirect calling mechanisms (see Section
\ref{sec:experiment_dynamicfeaturess}). As Figure
\ref{fig:output_candidate_explanation} illustrates, around 80 percent (except
for \sf{TimeClock}) all missed output candidates of each system were only
accessible through callback candidates, i.e., functions that are only
accessible through indirect calling mechanisms.

\section{Lessons Learned}
%\textcolor{blue}{\Blindtext}

\section{Related Work} \label{sec:related_work}
\cite{Hills:2013:ESP:2483760.2483786}
\cite{wassermann2007sound,minamide2005static}
%\textcolor{blue}{\blindtext\blindtext}

\section{Conclusion}

%\newpage
%References
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}